A Formal Model of Checked C

### 2023-04-07 sandp23/formal.tex
update formalism

\begin{figure}
  \small \centering
  \[ \hspace*{-1.5em}
\setlength{\arraycolsep}{4pt}
\begin{array}{l}
\begin{array}{ll}
       \text{Variables:}~ x
& \quad\text{Integers:}~n::=\mathbb{Z} 
\end{array}
\\[0.5em]
\begin{array}{llcllcl}
\text{Context Mode:} & m & ::= & \cmode \mid \umode \\
\text{Pointer Mode:} & \xi & ::= & m \mid \tmode \\
     \text{Type:}& \tau &::=& \tint\mid \tptr{\tau}{ \xi}
\\
\text{Expression:}& e & ::= & 
\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid\estar{e}\mid\eassign{e}{e}
  \\
&&\mid& \emalloc{\xi}{\omega} \mid \efree{e} \mid \elet{x}{e}{e}   \\
&&\mid& \ret{x}{\evalue{n}{\tau}}{e} \mid \eif{e}{e}{e} \\
&&\mid&\eunchecked{\overline{x}}{e}\mid \echecked{\overline{x}}{e}
\end{array}
    \end{array}
  \]
  \caption{\lang Syntax}
  \label{fig:checkc-syn}
\end{figure}

@@

\subsection{Syntax}\label{sec:syntax}
\Cref{fig:checkc-syn} shows the syntax of~\lang, 
which is based on a subset of the Compcert formalism (C formalism) \cite{Blazy2009}, with some features coming from the \checkedc formalism \cite{li22checkedc}.
We classify types as integers or pointers.
Every pointer type ($\tptr{\tau}{\xi}$) includes a
pointer mode annotation ($\xi$, the difference between context and pointer modes
is introduced shortly below) that is either checked (\cmode), tainted (\tmode),
or unchecked (\umode), and a content type ($\tau$) denoting the valid type it is pointed to.

\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$), 
pointer dereference ($\estar{e}$) and assignment ($\eassign{e_1}{e_2}$),
 static casts ($\ecast{\tau}{e}$),
memory allocations ($\emalloc{\xi}{\omega}$) and free ($\efree{e}$), 
along with expressions that require special handling, such as, tainted blocks ($\eunchecked{\overline{x}}{e}$),
and checked blocks ($\echecked{\overline{x}}{e}$).

We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\omega}{\xi}$), enabling the use of fixed addresses as pointers.
For example, $\evalue{0}{\tptr{\omega}{\xi}}$ (for any $\xi$ and $\omega$) represents a $\enull$ pointer.
The heap allocation $\emalloc{\xi}{\omega}$ includes a mode flag $\xi$ for allocating memory in $\xi$ memory heap regions, i.e., heaps are separated into three different regions, classified by $\xi$.
The $\echeckedtext$ and $\euncheckedtext$ expressions are used to delimit \textit{code} regions,
where the former represents the $\cmode$ code region in a program that are trusted by users, and the latter represents non-trust $\umode$ code region.
To guarantee the clean separation property, we extend the C syntax to include $\echecked{\overline{x}}{e}$ and $\eunchecked{\overline{x}}{e}$ blocks,
where $\overline{x}$ represents all free variables that are allowed to communicate between instructions outside and inside of the block $e$, and it cannot contain checked pointers.
$\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.

@@

\myparagraph{Disallowing Unsafe Types}
The well-formedness of these types are presented in~\Cref{app:le}.
It prevents unsafe types from being constructed.
  Consider the type ~\code{_t_ptr<_Ptr<int>>}, which describes a tainted pointer of a checked pointer.
  This is not well-formed in~\lang{} because it potentially exposes the checked pointer addresses in a $\umode$ region when the unchecked ($\umode$) pointer is used. 
  Nevertheless, we can have a checked pointer whose element is a tainted pointers\mzs{,}\mzr{:}~\eg~\code{_ptr<_t_Ptr<int>>} is a valid type.

@@

\begin{figure}
{\small
$\hspace*{-1.5em}
    \begin{array}{l}
    \begin{array}{lll}
r & ::= & e \mid \enull \mid \ebounds\\
E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}
    \mid \eassign{\evalue{n}{\tau}}{E}\\[0.2em]
&&\mid \efree{E}\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\\[0.2em]
&& \mid \eif{E}{e}{e} \mid \eunchecked{\overline{x}}{E} \mid \echecked{\overline{x}}{E}

@@

\begin{DIFnomarkup}
\begin{figure*}[t]
{\footnotesize
$\textcolor{blue}{\text{Dereference Operations:}}\\$
  \begin{mathpar}
    \inferrule[T-Def]
              {\xi \leq m \\\Gamma \vdash_m e : \tptr{\tau}{\xi}}
              {\Gamma \vdash_m \estar{e} : \tau}

        \inferrule[S-Def]{\xi \neq \tmode \\ \heap(n)=\evalue{n_a}{\tptr{\tau_a}{\xi}} }
    {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}

@@

    \inferrule[S-DefNull]{}{(\varphi,\Theta,\heap,\estar{\evalue{0}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\enull)}
\qquad
    \inferrule[S-DefBoundA]{n \in \Theta}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
\qquad
    \inferrule[S-DefBoundB]{\heap(n)=\evalue{n_a}{\tptr{\tau_a}{\xi}}\\\xi\neq \xi'}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\xi'}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
  \end{mathpar}

$\textcolor{blue}{\text{Malloc and Free:}}\\$
  \begin{mathpar}
    \inferrule[T-Malloc]
              {\xi \leq m}
              {\Gamma \vdash_m \emalloc{\xi}{\tau} : \tptr{\tau}{\xi}}

    \inferrule[T-Free]
              {\xi \leq m \\\Gamma \vdash_m e : \tptr{\tau}{\xi}}
              {\Gamma \vdash_m \efree{e} : \tint}

  \inferrule[S-Malloc]{\mathtt{alloc}(\heap,\Theta,\xi,\tau)=(n,\heap')}
   { (\varphi,\Theta,\heap,\emalloc{\xi}{\tau}) \longrightarrow (\varphi,\Theta',\heap',\evalue{n}{\tptr{\tau}{\xi}})}

  \inferrule[S-Free]{\xi \neq \tmode}
   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\xi}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{0}{\tint})}

  \inferrule[S-FreeT]{}
   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\tmode}}}) \longrightarrow (\varphi,\{n\}\cup\Theta,\heap,\evalue{0}{\tint})}

@@

$\textcolor{blue}{\text{Checked and Unchecked Context Switches:}}\\$
  \begin{mathpar}
    \inferrule[T-Checked]
              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
                     \\\\\fv(e)\subseteq\overline{x}\\\Gamma;\Theta \vdash_c e : \tau}
              {\Gamma;\Theta \vdash_m \echecked{\overline{x}}{e} : \tau}

    \inferrule[T-Unchecked]
              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
                \\\\ \fv(e)\subseteq\overline{x}\\\Gamma;\Theta \vdash_u e : \tau}
              {\Gamma;\Theta \vdash_m \eunchecked{\overline{x}}{e} : \tau}

    \inferrule[S-Unchecked]{}{(\varphi,\heap,\eunchecked{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
  \end{mathpar}
}
% {\footnotesize
% \begin{center}
% $
% \begin{array}{l}
% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
% \\[0.2em]
% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
% \end{array}
% $
% \end{center}
% }
{\footnotesize
\begin{center}
$
\begin{array}{l}
\tmode \le \xi
\qquad
m \le m
\\[0.2em]
\cmode(\tint)=\texttt{false}
\qquad
\cmode(\tptr{\omega}{\cmode})=\texttt{true}
\qquad
\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
\end{array}
$
\end{center}
}
\caption{Selected typing (\rulelab{T}-$X$) and semantic (\rulelab{S}-$X$) rules.}
\label{fig:type-system-1}
\end{figure*}
\end{DIFnomarkup}

@@

As partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{T-}$X$),
each typing judgment has the form $\Gamma\vdash_m e : \tau$,
which states that in a type environment $\Gamma$ (mapping variables to
their types), expression $e$ will have type $\tau$ if evaluated
in context mode $m$, indicating that the code is in $m$ region.
The operational semantics for \lang is defined as a small-step
transition relation with the judgment $ (\varphi,\Theta,\heap,e)
\longrightarrow_m (\varphi',\Theta',\heap',r)$, as shown in \Cref{fig:c-context}.
 Here, $\varphi$ is a
\emph{stack} mapping from variables to values $\evalue{n}{\tau}$,
$\Theta$ is a set of $\tmode$ mode memory addresses that are freed at the point of executing $e$
and $\heap$ is a \emph{heap} mapping addresses (integer literals) to
values $\evalue{n}{\ptr{\tau}{\xi}}$; for both we ensure
$\fv(\tau)=\emptyset$.
In a heap value $\evalue{n}{\ptr{\tau}{\xi}}$,
$\tau$ represents the value's content type, while $\xi$ indicates the mode of the heap location;
we classify a heap location into $\cmode$, $\tmode$, and $\umode$ modes, corresponding to the three kinds of pointer modes.
The complete set of typing rules are provided in~\Cref{rem-type,sec:rem-semantics}.

While heap bindings can change, stack bindings are immutable---once
variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
be updated. 
%We can model mutable stack variables as pointers into the mutable heap.
As mentioned, value $\evalue{0}{\tptr{\tau}{\xi}}$
represents a $\enull$ pointer when $\tau$ is a pointer type.
Correspondingly, $\heap(0)$ should always be undefined.
% 
The relation steps to a \emph{result} $r$, which is either an
expression, a $\enull$ or $\ebounds$ failure, represent an expression right
  after the reduction, a null-pointer dereference or heap-location-not-existence access,
respectively.
% 
Such failures are a \emph{good} outcome; stuck states
(non-value expressions that cannot transition to a result $r$)
characterizing undefined behavior.
%The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the stepped redex within $e$ was in a $\cmode$ or $\umode$ region.

The rules for the main operational semantics
judgment \emph{evaluation} are given at the bottom of
Fig.~\ref{fig:c-context}.
The first rule takes an expression $e$, decomposes
it into an \emph{evaluation context} $E$ and a sub-expression $e'$
(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
$e$), and then evaluates $e'$ according to the \emph{computation}
  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
whose rules are given along with type rules in Fig.~\ref{fig:type-system-1} (labeled as \rulelab{S-}$X$), discussed
shortly.
The $\mode$ function in Fig.~\ref{fig:c-context}
determines the context mode, i.e., region, that the expression $e'$ locates based on the context $E$.
If the context hole locates in an $\euncheckedtext$, i.e., $E[\eunchecked{\Box}]$, the context region mode is $\umode$,
while if the hole locates in a $\echeckedtext$, i.e., $E[\echecked{\Box}]$, its context mode is $\cmode$.
In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an unchecked function \code{server_loop}.

@@

In \lang, Context modes $m$ appearing in a type rule determine the code region 
that permits pointer dereferences and value-assignments, which also depends on the pointer modes.
We define a three point lattice $\xi_1 \le \xi_2$ \footnote{In typing rule, the lattice is usually used as $\xi \le m$ as $m$ represents context modes.} to describe such permission, where $\tmode \le \xi$ and $m \le m$.
This means that a $\tmode$ pointer can be dereferenced and value-assigned in any region, while $\cmode$ and $\umode$ pointers can only perform such operations in $\cmode$ and $\umode$ regions, respectively.
\ignore{
In \lang, we also provide a static casting operation in \Cref{rem-type},

@@

\myparagraph{Pointer Dereference}
The type and semantic rules for pointer dereference (the first block in \Cref{fig:type-system-1})
reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($\xi \le m$).

Rules \rulelab{S-DefT} and \rulelab{S-Def} are for $\tmode$ mode and other mode pointer dereferences, respectively.
Operation $\heap(n)$ not only accesses the $n$-th address location,
but also performs a $\enull$ check because $0$ is undefined in all heaps.
The pointer type ($\evalue{n}{\tptr{\tau}{\tmode}}$) in a dereference operation must have the same mode
as the one labeled in the type in the heap value $(\evalue{n_a}{\tptr{\tau_a}{\xi}}$).
The heap value type $\tptr{\tau_a}{\xi}$ is not an additional structure existed in the heap structure,
but an annotation lives in the compiler and is compiled as a dynamic check to ensure that pointers only access heap locations with right modes. We explain the dynamic check generation shortly below.
In addition to the above checks in rule \rulelab{S-Def},
we perform checks in \rulelab{S-DefT} $n \not\in \Theta$ to ensure that location $n$ is not freed in the $\tmode$ heap region.
If the above dynamic checks are violated, rules \rulelab{S-DefNull}, \rulelab{S-DefBoundA}, and \rulelab{S-DefBoundB} are performed, meaning that the evaluation is transitioned into a runtime error that is correctly captured by the \systemname compiler.

\myparagraph{Malloc and Free}
The second block (\rulelab{T-Malloc}, \rulelab{T-Free},
\rulelab{S-Malloc}, \rulelab{S-Free}, \rulelab{S-FreeT}) in \Cref{fig:type-system-1} defines
the type and semantic rules for $\texttt{malloc}$ and $\texttt{free}$ operations.
Rules $\rulelab{T-Malloc}$ and $\rulelab{T-Free}$ make sure that $\cmode$ and $\umode$ mode pointers are allocated and freed only in the $\cmode$ and $\umode$ regions, respectively, while $\tmode$ mode pointers are allowed to allocate and free in both regions.
The $\mathtt{alloc}(\heap,\Theta,\xi,\tau)$ function in rule \rulelab{S-Malloc} allocates a new pointer in the $\xi$ heap region.
If $\xi$ is $\tmode$ and $\Theta$ is not empty, the function picks a memory space in $\Theta$ as the new pointer;
otherwise, the function picks the $|\heap(\xi)|$ as the new pointer and increments $\heap(\xi)$'s cardinality.
The new location is labeled with a type annotation $\tptr{\tau}{\xi}$, indicating that the location is $\xi$ mode.
Every time when a deference and assignment is applied on a $\tmode$ pointer, the \systemname compiler applies a dynamic check to verify if the location is labeled with $\tmode$.
We can think of the dynamic check as $n \in \heap(\tmode)$, checking if a location $n$ is in a set $\heap(\tmode)$ containing all the $\tmode$ addresses. When a $\texttt{malloc}$ function is called to create a new $\tmode$ pointer, we expand set $\heap(\tmode)$ to include the new address. See \Cref{sec:compilation}.
Rules \rulelab{S-FreeT} and \rulelab{S-Free} define the \texttt{free} operation for $\tmode$ mode and other mode pointers, respectively.
Rule \rulelab{S-Free} "forgets" about the freed pointer, while \rulelab{S-FreeT} puts the freed $\tmode$ mode pointer address in the $\Theta$ set.

>>>>

\begin{figure}
  \small \centering
  \[ \hspace*{-1.5em}
\setlength{\arraycolsep}{4pt}
\begin{array}{l}
\begin{array}{ll}
       \text{Variables:}~ x
& \quad\text{Integers:}~n::=\mathbb{Z} 
\end{array}
\\[0.5em]
\begin{array}{llcllcl}
\text{Mode:} & m & ::= & \cmode \mid \umode \\
     \text{Type:}& \tau &::=& \tint\mid \tptr{\tau}{ m}
\\
\text{Expression:}& e & ::= & 
\evalue{n}{\tau} \mid x \mid \ebinop{e}{e}\mid \ecast{\tau}{e} \mid\estar{e}\mid\eassign{e}{e}
  \\
&&\mid& \emalloc{m}{\omega} \mid \efree{e} \mid \elet{x}{e}{e}   \\
&&\mid& \ret{x}{\evalue{n}{\tau}}{e} \mid \eif{e}{e}{e} \\
&&\mid&\toa{m}{\overline{x}}{e}
\end{array}
    \end{array}
  \]
  \caption{\lang Syntax}
  \label{fig:checkc-syn}
\end{figure}

@@

\subsection{Syntax}\label{sec:syntax}
\Cref{fig:checkc-syn} shows the syntax of~\lang, 
which is based on a subset of the Compcert formalism (C formalism) \cite{Blazy2009}, with some features coming from the \checkedc formalism \cite{li22checkedc}.
We classify types as integers or pointers.
Every pointer type ($\tptr{\tau}{m}$) includes a
 mode annotation that is either tainted (\umode) and non-tainted (\cmode),
and a content type ($\tau$) denoting the valid type it is pointed to.

\lang expressions include common expressions such as addition ($\ebinop{e_1}{e_2}$), 
pointer dereference ($\estar{e}$) and assignment ($\eassign{e_1}{e_2}$),
 static casts ($\ecast{\tau}{e}$),
memory allocations ($\emalloc{m}{\tau}$) and free ($\efree{e}$), 
along with expressions that require special handling, such as, context switching blocks ($\toa{m}{\overline{x}}{e}$).

We denote integer literals $n$ with a type $\tau$ (\ie $\tint$ or $\tptr{\tau}{m}$), enabling the use of fixed addresses as pointers.
For example, $\evalue{0}{\tptr{\tau}{m}}$ (for any $m$ and $\tau$) represents a $\enull$ pointer.
The heap allocation $\emalloc{m}{\tau}$ includes a mode flag $m$ for allocating memory in $m$ memory heap regions, i.e., heaps are separated into two different regions, classified by $m$.
Expression $\toa{m}{\overline{x}}{e}$ is used to delimit \textit{code} regions, where $m$ indicates that $e$ locates in the $m$ code region.
To guarantee the clean separation property, we extend the C syntax to include $\toa{m}{\overline{x}}{e}$ blocks,
which represents the flag assigning to a function to indicate if the function is tainted or not.
$\overline{x}$ represents all free variables that are allowed to communicate between instructions outside and inside of the block $e$, and it cannot contain $\cmode$ mode pointers.
$\erettext$ is introduced by the semantics when evaluating a \texttt{let} binding; explained shortly below.

@@

\myparagraph{Disallowing Unsafe Types}
The well-formedness of these types are presented in~\Cref{app:le}.
It prevents unsafe types from being constructed.
  Consider the type ~\code{_t_ptr<_Ptr<int>>}, which describes a tainted pointer of a $\cmode$ mode pointer.
  This is not well-formed in~\lang{} because it potentially exposes the $\cmode$ mode pointer addresses in a $\umode$ region when the tainted ($\umode$) pointer is used. 
  Nevertheless, we can have a $\cmode$ mode pointer whose element is a tainted pointers\mzs{,}\mzr{:}~\eg~\code{_ptr<_t_Ptr<int>>} is a valid type.

@@

\begin{figure}
{\small
$\hspace*{-1.5em}
    \begin{array}{l}
    \begin{array}{lll}
r & ::= & e \mid \enull \mid \ebounds\\
E &::=& \Box \mid \ebinop{E}{e} \mid \ebinop{\evalue{n}{\tau}}{E}\mid \ecast{\tau}{E} \mid\estar{E}\mid\eassign{E}{e}
    \mid \eassign{\evalue{n}{\tau}}{E}\\[0.2em]
&&\mid \efree{E}\mid \elet{x}{E}{e}\mid\ret{x}{\evalue{n}{\tau}}{E}\\[0.2em]
&& \mid \eif{E}{e}{e} \mid \toa{m}{\overline{x}}{E}

@@

\begin{DIFnomarkup}
\begin{figure*}[t]
{\footnotesize
$\textcolor{blue}{\text{Dereference Operations:}}\\$
  \begin{mathpar}
    \inferrule[T-Def]
              {m' \leq m \\\Gamma \vdash_m e : \tptr{\tau}{m'}}
              {\Gamma \vdash_m \estar{e} : \tau}

        \inferrule[S-Def]{ \heap(n)=\evalue{n_a}{\tptr{\tau_a}{\umode}} }
    {(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\umode}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{n_a}{\tau})}

@@

    \inferrule[S-DefNull]{}{(\varphi,\Theta,\heap,\estar{\evalue{0}{\tptr{\tau}{m}}}) \longrightarrow (\varphi,\Theta,\heap,\enull)}
\qquad
    \inferrule[S-DefBoundA]{n \in \Theta}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{\umode}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
\qquad
    \inferrule[S-DefBoundB]{\heap(n)=\evalue{n_a}{\tptr{\tau_a}{m}}\\ m\neq m'}{(\varphi,\Theta,\heap,\estar{\evalue{n}{\tptr{\tau}{m'}}}) \longrightarrow (\varphi,\Theta,\heap,\ebounds)}
  \end{mathpar}

$\textcolor{blue}{\text{Malloc and Free:}}\\$
  \begin{mathpar}
    \inferrule[T-Malloc]
              {m' \leq m}
              {\Gamma \vdash_m \emalloc{m'}{\tau} : \tptr{\tau}{m'}}

    \inferrule[T-Free]
              {m' \leq m \\\Gamma \vdash_m e : \tptr{\tau}{m'}}
              {\Gamma \vdash_m \efree{e} : \tint}

  \inferrule[S-Malloc]{\mathtt{alloc}(\heap,\Theta,m,\tau)=(n,\heap')}
   { (\varphi,\Theta,\heap,\emalloc{m}{\tau}) \longrightarrow (\varphi,\Theta',\heap',\evalue{n}{\tptr{\tau}{m}})}

  \inferrule[S-Free]{}
   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\cmode}}}) \longrightarrow (\varphi,\Theta,\heap,\evalue{0}{\tint})}

  \inferrule[S-FreeT]{}
   { (\varphi,\Theta,\heap,\efree{\evalue{n}{\tptr{\tau}{\umode}}}) \longrightarrow (\varphi,\{n\}\cup\Theta,\heap,\evalue{0}{\tint})}

@@

$\textcolor{blue}{\text{Context Switches:}}\\$
  \begin{mathpar}
    \inferrule[T-Region]
              {\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))\\\neg\cmode(\tau)
                     \\\fv(e)\subseteq\overline{x}\\\Gamma;\Theta \vdash_{m'} e : \tau}
              {\Gamma;\Theta \vdash_m \toa{m'}{\overline{x}}{e} : \tau}

    \inferrule[S-Region]{}{(\varphi,\heap,\toa{m}{\overline{x}}{\evalue{n}{\tau}}) \longrightarrow (\varphi,\heap,\evalue{n}{\tau})}
  \end{mathpar}
}
% {\footnotesize
% \begin{center}
% $
% \begin{array}{l}
% \fm(e)\triangleq(\exists x\; n\; \tau. e=x+\evalue{n}{\tau}) \vee (\exists n\;\tau. e = \evalue{n}{\tau})
% \\[0.2em]
% \tau[\overline{e} / \overline{x}]\texttt{(with types }\evalue{\overline{x}}{\overline{\tau}}\texttt{)}\triangleq \forall e_i\in\overline{e}\;x_i\in\overline{x}\;\tau_i\in\overline{\tau}\;.\;\tau_i = \tint \wedge (x_i \in \fv(\tau) \Rightarrow \fm(e_i)) \Rightarrow \tau[e_i / x_i]
% \end{array}
% $
% \end{center}
% }
{\footnotesize
\begin{center}
$
\begin{array}{l}
\umode \le \cmode
\qquad
m \le m
\\[0.2em]
\cmode(\tint)=\texttt{false}
\qquad
\cmode(\tptr{\omega}{\cmode})=\texttt{true}
\qquad
\cmode(\tptr{\omega}{\xi})=\texttt{false}\;\;{[\emph{owise}]}
\end{array}
$
\end{center}
}
\caption{Selected typing (\rulelab{T}-$X$) and semantic (\rulelab{S}-$X$) rules.}
\label{fig:type-system-1}
\end{figure*}
\end{DIFnomarkup}

@@

As partly shown in \Cref{fig:type-system-1} (labeled as \rulelab{T-}$X$),
each typing judgment has the form $\Gamma\vdash_m e : \tau$,
which states that in a type environment $\Gamma$ (mapping variables to
their types), expression $e$ will have type $\tau$ if evaluated
in context mode $m$, indicating that the code is in $m$ region.
The operational semantics for \lang is defined as a small-step
transition relation with the judgment $ (\varphi,\Theta,\heap,e)
\longrightarrow_m (\varphi',\Theta',\heap',r)$, as shown in \Cref{fig:c-context}.
 Here, $\varphi$ is a
\emph{stack} mapping from variables to values $\evalue{n}{\tau}$,
$\Theta$ is a set of tainted ($\umode$ mode) memory addresses that are freed at the point of executing $e$
and $\heap$ is a \emph{heap} mapping addresses (integer literals) to
values $\evalue{n}{\ptr{\tau}{m}}$; for both we ensure
$\fv(\tau)=\emptyset$.
In a heap value $\evalue{n}{\ptr{\tau}{m}}$,
$\tau$ represents the value's content type, while $m$ indicates the mode of the heap location;
we classify a heap location into $\cmode$ and $\umode$ modes, corresponding to the two kinds of pointer modes.
The complete set of typing rules are provided in~\Cref{rem-type,sec:rem-semantics}.

While heap bindings can change, stack bindings are immutable---once
variable $x$ is bound to $\evalue{n}{\tau}$ in $\varphi$, that binding will not
be updated. 
%We can model mutable stack variables as pointers into the mutable heap.
As mentioned, value $\evalue{0}{\tptr{\tau}{m}}$
represents a $\enull$ pointer when $\tau$ is a pointer type.
Correspondingly, $\heap(0)$ should always be undefined.
% 
The relation steps to a \emph{result} $r$, which is either an
expression, a $\enull$ or $\ebounds$ failure, represent an expression right
  after the reduction, a null-pointer dereference or heap-location-not-existence access,
respectively.
% 
Such failures are a \emph{good} outcome; stuck states
(non-value expressions that cannot transition to a result $r$)
characterizing undefined behavior.
%The context mode $m$ (in $\longrightarrow_{m}$) indicates whether the stepped redex within $e$ was in a $\cmode$ or $\umode$ region.

The rules for the main operational semantics
judgment \emph{evaluation} are given at the bottom of
Fig.~\ref{fig:c-context}.
The first rule takes an expression $e$, decomposes
it into an \emph{evaluation context} $E$ and a sub-expression $e'$
(such that replacing the hole $\Box$ in $E$ with $e'$ would yield
$e$), and then evaluates $e'$ according to the \emph{computation}
  relation $(\varphi,\heap,e') \longrightarrow (\varphi,\heap,e'')$,
whose rules are given along with type rules in Fig.~\ref{fig:type-system-1} (labeled as \rulelab{S-}$X$), discussed
shortly.
The $\mode$ function in Fig.~\ref{fig:c-context}
determines the context mode, i.e., region, that the expression $e'$ locates based on the context $E$.
If the context hole locates in an $\totext(m)$, i.e., $E[\toa{m}{\overline{x}}{\Box}]$, the context region mode is $m$.
In \Cref{lst:humantaint}, the function call \code{handle_request} is in $\umode$ region since it is inside an tainted function \code{server_loop}.

@@

In \lang, modes $m$ appearing in a type rule determine the code region 
that permits pointer dereferences and value-assignments, which also depends on the pointer modes.
We define a two point lattice $m_1 \le m_2$, with $\umode \le \cmode$, to describe such permission.
This means that a $\umode$ mode pointer can be dereferenced and value-assigned in any region, while $\cmode$ mode pointers can only perform such operations in $\cmode$ region.
\ignore{
In \lang, we also provide a static casting operation in \Cref{rem-type},

@@

\myparagraph{Pointer Dereference}
The type and semantic rules for pointer dereference (the first block in \Cref{fig:type-system-1})
reflect the key \lang feature, where our type checker directs the insertions of dynamic checks executed in the evaluation stage.
The type rule (\textsc{T-Def}) ensures that pointers are used with the right modes in the right region ($m' \le m$).

Rules \rulelab{S-DefT} and \rulelab{S-Def} are for $\umode$ mode and other mode pointer dereferences, respectively.
Operation $\heap(n)$ not only accesses the $n$-th address location,
but also performs a $\enull$ check because $0$ is undefined in all heaps.
The pointer type ($\evalue{n}{\tptr{\tau}{\umode}}$) in a dereference operation must have the same mode
as the one labeled in the type in the heap value $(\evalue{n_a}{\tptr{\tau_a}{m}}$).
The heap value type $\tptr{\tau_a}{m}$ is not an additional structure existed in the heap structure,
but an annotation lives in the compiler and is compiled as a dynamic check to ensure that pointers only access heap locations with right modes. We explain the dynamic check generation shortly below.
In addition to the above checks in rule \rulelab{S-Def},
we perform checks in \rulelab{S-DefT} $n \not\in \Theta$ to ensure that location $n$ is not freed in the $\umode$ heap region.
If the above dynamic checks are violated, rules \rulelab{S-DefNull}, \rulelab{S-DefBoundA}, and \rulelab{S-DefBoundB} are performed, meaning that the evaluation is transitioned into a runtime error that is correctly captured by the \systemname compiler.

\myparagraph{Malloc and Free}
The second block (\rulelab{T-Malloc}, \rulelab{T-Free},
\rulelab{S-Malloc}, \rulelab{S-Free}, \rulelab{S-FreeT}) in \Cref{fig:type-system-1} defines
the type and semantic rules for $\texttt{malloc}$ and $\texttt{free}$ operations.
Rules $\rulelab{T-Malloc}$ and $\rulelab{T-Free}$ make sure that $\cmode$ mode pointers are allocated and freed only in the $\cmode$ region, while $\umode$ mode pointers are allowed to allocate and free in both regions.
The $\mathtt{alloc}(\heap,\Theta,m,\tau)$ function in rule \rulelab{S-Malloc} allocates a new pointer in the $m$ heap region.
If $m$ is $\umode$ and $\Theta$ is not empty, the function picks a memory space in $\Theta$ as the new pointer;
otherwise, the function picks the $|\heap|$ as the new pointer and increments $\heap$'s cardinality.
The new location is labeled with a type annotation $\tptr{\tau}{m}$, indicating that the location is $m$ mode.
Every time when a deference and assignment is applied on a $\umode$ pointer, the \systemname compiler applies a dynamic check to verify if the location is labeled with $\umode$.
We can think of the dynamic check as $n \in \heap(\umode)$, checking if a location $n$ is in a set $\heap(\umode)$ containing all the $\umode$ addresses. When a $\texttt{malloc}$ function is called to create a new $\umode$ pointer, we expand set $\heap(\umode)$ to include the new address. See \Cref{sec:compilation}.
Rules \rulelab{S-FreeT} and \rulelab{S-Free} define the \texttt{free} operation for $\umode$ mode and other mode pointers, respectively.
Rule \rulelab{S-Free} "forgets" about the freed pointer, while \rulelab{S-FreeT} puts the freed $\umode$ mode pointer address in the $\Theta$ set.

\myparagraph{Context Switching Blocks}
The execution of a $\totext$ block represents 
the context switching from a $\cmode$ to an $\umode$ region, or vice versa,
with its type and semantic rules given in \Cref{fig:type-system-1}.
In this context switching, to guarantee the $\cmode$ mode pointer non-exposure property, i.e.,
$\cmode$ pointers are not allowed to go cross different regions, which is guaranteed by the predicates 
$\forall x\in\overline{x}\;.\;\neg\cmode(\Gamma(x))$ and $\neg\cmode(\tau)$ \footnote{$\cmode(-)$ checks the mode of a pointer.},
as well as the check that all free variables in the block content $e$ are in $\overline{x}$ ($\fv(e)\subseteq\overline{x}$).
For example, \code{StringAuth} in \Cref{subsub:gencregion} is a trampoline function that disallows
$\cmode$ pointers as arguments and return values.
The use of the function in the following \code{_T_StringAuth}, which is in $\umode$ region,
cannot legally acknowledges any $\cmode$ pointers; otherwise, we might expose a $\cmode$ pointer address to unsafe code regions.
In \lang, we actually permits the accesses of $\cmode$ pointers inside \code{StringAuth},
since the function body of a trampoline function is in $\cmode$ region.
More information is given in \Cref{subsub:gencregion}.
